---
title: "Advanced R Notebook"
output: html_notebook
---

Dejan Sarka - presenter

#Introduction to Data Science
reinforcement learning

https://goo.gl/K9YovT

What is Data Science? Extracting knowledge from data - statistics and programming
Statistics is a science - data mining and machine learning are more business oriented

Statistics  small <30, medium, >200 < 100K) Large > 100K
DM - medium
ML - medium and large

##Vectors
ordered collections of numbers
created with c() to concatenate elements or sub-vectors
rep() to repeat elements or patterns
seq() or m:n to generate sequences

functions can be applied to vectors without loops
use the [] operator to select elements

##Packages
packages are stored in library
installed.package = installed packages
library() list packages in library


```{r basic expressions and variables}

1+1
2 + 3 * 4
3 ^ 3
sqrt(81)
pi

#Check built-in constants
??"constants"

#Sequences
rep(1, 5)
seq(4, 8)
seq(4, 20, by = 3)

#Assignment
x <- 2
y <- 3
z <- 4
X + y + z #error as variables are case sensitive
x + y + z

#Vectors
x <- c(2, 0, 0, 4); x
assign("y", c(1, 9, 9, 9)); y
c(5, 4, 3, 2) >- z; z
q = c(1, 2, 3, 4); q

installed.packages()
.libPaths()
install.packages("RODBC")
library(RODBC)

# # Ad-hoc connection
# con <- odbcDriverConnect('driver={SQL Server};server=SQL2017EIM;
#                          database=AdventureWorksDW2017;uid=RUser;pwd=Pa$$w0rd')
# 
# # DSN connection
# con <- odbcConnect("AWDW", uid = "RUser", pwd = "Pa$$w0rd")
# 
# # Read SQL Server data
# sqlQuery(con,
#          "SELECT CustomerKey,
#          EnglishEducation AS Education,
#          Age, NumberCarsOwned, BikeBuyer
#          FROM dbo.vTargetMail;")
# close(con)
```

##Collections and Objects
###Matrices (Arrays) 
  -multidimensional generalizations of vectors
  
###Factors 
  -compact categorical data handling
  -Can have discrete or coninuous values
  -Discrete values can be nominal or catgorical when labeling, or ordinal if there is intrinsic order to data
  -Levels represent distinct values - order must be defined
  
###Lists 
  - ordered collection of different data structures
  
###Data Frames or matrix-like structures
  

-Functions


```{r}
# Matrix
x = c(1, 2, 3, 4, 5, 6); x
Y = array(x, dim = c(2, 3)); Y
Z = matrix(x, 2, 3, byrow = F); Z
U = matrix(x, 2, 3, byrow = T); U

# Using explicit names
rnames = c("rrr1", "rrr2")
cnames = c("ccc1", "ccc2", "ccc3")
V = matrix(x, 2, 3, byrow = T,
           dimnames = list(rnames, cnames))
V

# Array
rnames = c("rrr1", "rrr2")
cnames = c("ccc1", "ccc2", "ccc3")
pnames = c("ppp1", "ppp2", "ppp3")
Y = array(1:18, dim = c(2, 3, 3),
          dimnames = list(rnames, cnames, pnames))
Y

# Factor
x = c("good", "moderate", "good", "bad", "bad", "good"); x
y = factor(x); y
z = factor(x, order = TRUE); z
w = factor(x, order = TRUE,
           levels = c("bad", "moderate", "good")); w

# List
L = list(name1 = "ABC", name2 = "DEF",
         no.children = 2, children.ages = c(3, 6))
L
L[[1]]
L[[4]]
L[[4]][2]

# Data frame
CategoryId = c(1, 2, 3, 4)
CategoryName = c("Bikes", "Components", "Clothing", "Accessories")
ProductCategories = data.frame(CategoryId, CategoryName)
ProductCategories

library(RODBC)
# Reading in a data frame from SQL Server
# con <- odbcConnect("AWDW", uid = "RUser", pwd = "Pa$$w0rd")
# TM <-
#   sqlQuery(con,
#            "SELECT CustomerKey,
#             EnglishEducation AS Education,
#             Age, NumberCarsOwned, BikeBuyer
#           FROM dbo.vTargetMail;")
# close(con)
# head(TM, 5)
# tail(TM, 5)
# TM[1:5, 1:5]
# TM[1:5, c("CustomerKey", "BikeBuyer")]
# TM[TM$CustomerKey >= 29479, c("CustomerKey", "BikeBuyer")]

# # Check the complete data frame
# View(TM)
# 
# # Crosstabulation of BikeBuyer and NumberCarsOwned
# table(TM$NumberCarsOwned, TM$BikeBuyer)
# 
# # Data table
# # install.packages("data.table")
# library(data.table)
# 
# dtTM <- data.table(TM)
# class(dtTM)
# # See explicit row numbers
# dtTM
# 
# # View works
# View(dtTM)
# 
# # Aggregation
# dtTM[, mean(Age)]
# # The following line does not work
# TM[, mean(Age)]
# 
# # More aggregations
# dtTM[, .N]   # counts
# dtTM[, .(Count=.N), by = NumberCarsOwned]
# dtTM[, .(AverageAge=mean(Age)), by = NumberCarsOwned]
# 
# # Filter
# dtTM[NumberCarsOwned > 1, .(AverageAge=mean(Age)),
#      by = NumberCarsOwned]
# 
# # Order
# head(dtTM[order(-CustomerKey)], 5)

```


#Data Preparation and Overview
